On the properties of discs around accreting brown dwarfs

### 2010-05-26 BD_disc/paper.tex
inner edge stuff ringfenced
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2287 25464f21-6099-402c-b30a-640f1d40d38b

**GOT HERE NEED NEW FIGURE FOR INNER EDGE AND NUMBERS WRONG***
CHANGE flareinner.ps and flareinnermag.ps

@@

Figure \ref{flare_inner} shows the total SEDs for the BDD system of
Figures \ref{flare_-12_183} and \ref{flare_-7_193} as the top and
bottom panels respectively. The left panels show the flux over all the
ten inclinations (see Table \ref{par_space_table}), with the dashed
line showing the inclination at which a sharp fall in flux is
seen. This inclination will be the angle at which the star and inner
disc becomes obscured by the flared outer disc. These inclinations are
71$^{\circ}$ and 56$^{\circ}$ for \logmdot = $-$12 and $-$7
respectively. This effectively means that for higher accretion rates a
more significant fraction of the stars, in a given population, may
have flux below a threshold detection limit. The right panel of Figure
\ref{flare_inner} shows the angles of 0, 27, 39, and 48 $^{\circ}$,
close to face on. Comparing the top and bottom panel show that flux
levels for the higher accretion rate, with the larger and curved inner
edge, vary less distinctly with inclination angle than the systems
with a vertical wall. These trends although demonstrated for one
subset are apparent throughout the grid, i.e. for each mass and age.

\begin{figure*}
  \centering
  \includegraphics[scale=0.6,angle=90]{./Fig/flare_inner.ps}
  \caption{Figure showing both total SEDs of the systems shown in
    Figure \ref{flare_-12_183} and \ref{flare_-7_193} as the top and
    bottom panels respectively. The left panel shows all inclinations
    (see Table \ref{par_space_table}), with the obscuration angle
    shown as a dashed line, 71$^{\circ}$ and 56$^{\circ}$ for
    \logmdot = $-$12 and $-$7 respectively. The right panels show only the
    inclinations close to face on, i.e. 0, 27, 39, and 48 $^{\circ}$,
    with a clear dependence on inclination of the flux levels for the
    curved inner wall (lower right panel).\label{flare_inner}}
\end{figure*}

The changes in flux with inclination will clearly lead to changes in
the observed magnitudes and colours with inclination. Figure
\ref{flare_inner_mag} shows the $M_V$ and $M_J$ magnitudes and the
$(J-K)_0$ colour index, in the top, middle and bottom panels
respectively. The colours and magnitudes are then marked as crosses as
a function of inclination.  The systems shown in all panels have an
age of 1 Myr, mass of $0.04M_{\odot}$, rotation rates of 0.5 and an
areal coverage of 10\%. The systems with accretion rates of \logmdot =
$-$12, $-$9 and $-$7 are shown in black, red and blue, respectively,
with the lowest panel omitting \logmdot = $-$9, and restricting the
inclination to viewing angles close to face-on.  The top and middle
panels also show the system with the highest accretion rate but with a
rotation rate of 5 days as a dashed line. The vertical dotted lines in
the lowest panel simply illustrate the total change in magnitude, over
the shown inclination range.

\begin{figure*}
  \centering
  \includegraphics[scale=0.6,angle=90]{./Fig/flare_inner_mag.ps}
  \caption{Figure showing $M_V$, $M_J$ magnitudes and $(J-K)_0$ colour
    index as a function of inclination, shown as top, middle and lower
    panels respectively. The magnitude and colours of systems with
    $M_*=0.04M_{\odot}$, an age of 1 Myr, areal coverage of 10\% and
    rotation rate of 0.5 days are shown as crosses. The systems with
    accretion rates of \logmdot = $-$12, $-$9 and $-$7, are shown in black,
    red and blue respectively. The lowest panel omits the \logmdot $-$9
    systems. The top and middle panels also the systems with the
    highest accretion rate but for a rotation rate of 5 days. The
    lowest panel also shows, as horizontal dotted lines, the total
    change of colour with the illustrated
    inclinations. \label{flare_inner_mag}}
\end{figure*}

Accretion flux was shown to dominate the underlying, intrinsic,
photospheric SED for accretion rates of \logmdot = $-$9 (see Figure
\ref{acc_dom}). As one would expect this leads to significant changes
in the derived photometric magnitudes for bands blueward of a few
microns, where the accretion and photospheric flux dominate. This is
shown in the top panel of Figure \ref{flare_inner_mag}. As the
accretion rate increases the system moves to brighter magnitudes in
$M_V$. The move to brighter magnitudes becomes significant once the
accretion rate exceeds \logmdot = $-$9. Again moving from the fast to slow
rotation period increases the potential energy released and in this
case results in a brighter $M_V$ magnitude. As one moves to longer
wavelength photometric bands, as shown in the middle panel. The
increase in accretion rate affects these magnitudes, i.e. $M_J$ are
les significantly affected. This is due to the significance of the
inner edge at these longer wavelengths. The lowest panel shows that as
the accretion rate is increase the IR colours generally get
redder. Again, this is caused by interaction with the disc, as more
flux is reprocessed by the inner edge.

The top and middle panels of Figure \ref{flare_inner_mag} shows the
change in magnitude caused by obscuration from the flared outer
disc. The occultation starts earlier for the higher accretion rates,
with the magnitudes in $M_V$ and $M_J$ moving fainter at earlier
inclinations for the higher accretion rates.

XXXNOT SURE ABOUT THIS NEXT BIT, PERHAPS A LITTLE WEAKXXX
**SENSE OF ARGUMENT RIGHT BUT NUMBERS WRONG***

The lowest panel of Figure \ref{flare_inner_mag} shows the difference
in the inclination dependence of the IR colour for large curved
(\logmdot = $-$7) and smaller vertical (\logmdot = $-$12) inner
walls. The $(J-K)_0$ colour of the higher accretion rate system
changes less as a function of inclination when compared to the lower
accretion rate system. A total change in $(J-K)_0$ of $\approx$ 0.05
and 0.1 are found for \logmdot = $-$12 and $-$7 respectively. This
halving in the change in IR colour as a function of inclination for
viewing angles close to face-on, is caused by the larger (vertically)
and convex inner wall in the extreme accreting case. For the convex
inner wall the effective emitting area one observes does not change as
much as a vertical wall as a function of inclination, leading to this
reduced dependence of IR colour on inclination.

@@

This scatter for both typical and extreme accretors is a strong
function of inclination, where, as the inclination is increased the
objects are pushed lower in the CMD. Indeed, for the edge on cases
some objects have magnitudes fainter than those shown (for instance
$M_V\approx 20$). This is expected as the star becomes obscured by the
flared disc, interestingly for typical systems the bottom right panels
show that this occurs for inclinations above around 71$^{\circ}$ (as
found in Figure \ref{flare_inner}) in most cases.  However, even for
the lower inclination angles some objects have very faint magnitudes,
this is due to the disc flaring leading to a smaller opening angle and
is discussed in Section \ref{disc_struct}. Crucially, the top panels
show that the scatter from the isochrone is, generally, correlated
with accretion rate. Effectively, as the accretion rate increases the
BDD system moves farther away from the isochrone and is therefore less
likely to be classified as a BDD system and included in any target
samples of such objects. Overall, the dominant scattering effect for
BDD systems in an optical CMD appears to be caused by accretion rate
and inclination, and therefore obscuration effects of the disc on the
star. This suggests that for a given photometric survey of BDD systems
accreting at typical accretion rates and with an expected range of
inclinations (centred on around 60$^{\circ}$) one would expect to
exclude a significant fraction of these objects from any isochrone
based selection.

@@

Subsequently, the SEDs of BDD systems with typical accretion rates and
associated discs, are changed significantly from the assumed
underlying photospheric model flux, and therefore become difficult to
classify. In Section \ref{observables} we have shown that the BD
photosphere becomes veiled by the accretion flux for rates of
\logmdot$>-$9. The outer disc flaring observed in the BDD systems was
shown to cause occultation and a subsequent, sharp, fall in flux at an
inclination which decreases for more systems with higher accretion
rates. We have also shown that for extreme accretion rates the inner
wall increases in size and becomes convex in shape. We have also shown
that this leads to a reducedq dependence on inclination (at
inclinations close to face-on) of the IR colours, compared to a
vertical wall.

>>>>

@@

Figure \ref{flare_inc} shows the total SEDs for the BDD system of
Figures \ref{flare_-12_183} and \ref{flare_-7_193} as the top and
bottom panels respectively. The lines show the flux over all the ten
inclinations (see Table \ref{par_space_table}), with the dashed line
showing the inclination at which a sharp fall in flux is seen. This
inclination will be the angle at which the star and inner disc becomes
obscured by the flared outer disc. These inclinations are 71$^{\circ}$
and 56$^{\circ}$ for \logmdot = $-$12 and $-$7 respectively. This
effectively means that for higher accretion rates a more significant
fraction of the stars, in a given population, may have flux below a
threshold detection limit.

\begin{figure*}
  \centering
  \includegraphics[scale=0.6,angle=90]{./Fig/flare_inc.ps}
  \caption{Figure showing both total SEDs of the systems shown in
    Figure \ref{flare_-12_183} and \ref{flare_-7_193} as the top and
    bottom panels respectively. The lines show the SEDs of all
    inclinations (see Table \ref{par_space_table}), with the
    obscuration angle shown as a dashed line, 71$^{\circ}$ and
    56$^{\circ}$ for \logmdot = $-$12 and $-$7
    respectively.\label{flare_inc}}
\end{figure*}

The changes in flux with inclination will clearly lead to changes in
the observed magnitudes and colours with inclination. Figure
\ref{flare_inc_mag} shows the $M_V$ and $M_J$ magnitudes in the top
and bottom panels respectively. The magnitudes are then marked as
crosses as a function of inclination. The systems shown in both panels
have an age of 1 Myr, mass of $0.04M_{\odot}$, rotation rates of 0.5
and an areal coverage of 10\%. The systems with accretion rates of
\logmdot = $-$12, $-$9 and $-$7 are shown in black, red and blue,
respectively. The panels also show the system with the highest
accretion rate but with a rotation rate of 5 days as a dashed
line. The vertical dotted lines in the lowest panel simply illustrate
the total change in magnitude, over the shown inclination range.

\begin{figure*}
  \centering
  \includegraphics[scale=0.6,angle=90]{./Fig/flare_inc_mag.ps}
  \caption{Figure showing $M_V$ and $M_J$ magnitudes as a function of
    inclination, shown as top, middle and lower panels
    respectively. The magnitudes of systems with $M_*=0.04M_{\odot}$,
    an age of 1 Myr, areal coverage of 10\% and rotation rate of 0.5
    days are shown as crosses. The systems with accretion rates of
    \logmdot = $-$12, $-$9 and $-$7, are shown in black, red and blue
    respectively. The panels also the systems with the highest
    accretion rate but for a rotation rate of 5 days. The lowest panel
    also shows, as horizontal dotted lines, the total change of colour
    with the illustrated inclinations. \label{flare_inc_mag}}
\end{figure*}

Accretion flux was shown to dominate the underlying, intrinsic,
photospheric SED for accretion rates of \logmdot = $-$9 (see Figure
\ref{acc_dom}). As one would expect this leads to significant changes
in the derived photometric magnitudes for bands blueward of a few
microns, where the accretion and photospheric flux dominate. This is
shown in the top panel of Figure \ref{flare_inc_mag}. As the accretion
rate increases the system moves to brighter magnitudes in $M_V$. The
move to brighter magnitudes becomes significant once the accretion
rate exceeds \logmdot = $-$9. Again moving from the fast to slow
rotation period increases the potential energy released and in this
case results in a brighter $M_V$ magnitude. As one moves to longer
wavelength photometric bands, as shown in the middle panel. The
increase in accretion rate affects these magnitudes, i.e. $M_J$ are
less significantly affected. Figure \ref{flare_inc_mag} also shows the
change in magnitude caused by obscuration from the flared outer
disc. The occultation starts earlier for the higher accretion rates,
with the magnitudes in $M_V$ and $M_J$ moving fainter at earlier
inclinations for the higher accretion rates.

XXXTHIS SECTION IS ON THE INNER EDGE AND NEEDS SORTINGXXX
179 
221

Figure \ref{inner_inc} shows the SEDs for the viewing angles of 0, 27,
39, and 48 $^{\circ}$, close to face on. The top and bottom panels of
Figure \ref{inner_inc} show SEDs for models with a curved, as
exemplified by Figure \ref{flare_-7_193}, and flat inner wall, as
shown in Figure \ref{flare_-12_183} respectively. The specific models
chosen have been selected to minimise flux differences between the two
SEDs but have distinct inner edges. XXXMODELSXXX Comparing the top and
bottom panel show that flux levels for the higher accretion rate, with
the larger and curved inner edge, vary less distinctly with
inclination angle than the systems with a vertical wall.

\begin{figure*}
  \centering
  \includegraphics[scale=0.6,angle=90]{./Fig/inner_inc.ps}
  \caption{Figure showing SEDs of systems with a curved and flat inner
    wall as the top and bottom panels respectively. The lines show the
    SEDs at viewing angles of 0, 27, 39, and 48 $^{\circ}$, close to
    face-on.\label{inner_inc}}
\end{figure*}

Figure \ref{inner_inc} shows only that for inclinations close to
face-on, the curved inner edge has SEDs with a stronger dependence on
the viewing angle compared to the model with a vertical inner
wall. This leads to a greater dependence of IR colour with
inclination.  Figure \ref{inner_mag} shows the difference in the
inclination dependence of the IR colour for large curved and smaller
vertical inner walls, for the systems shown in Figure
\ref{inner_inc}.

\begin{figure*}
  \centering
  \includegraphics[scale=0.6,angle=90]{./Fig/inner_mag.ps}
  \caption{Figure showing the $(J-K)_0$ colour against the
    inclinations of 0, 27, 39, and 48 $^{\circ}$, for the models shown
    in Figure \ref{inner_inc}. The models with the flat, vertical
    inner wall and curved inner edge are shown as the solid and dashed
    lines respectively. The horizontal dotted lines show the change in
    the colour index.\label{inner_mag}}
\end{figure*}

The $(J-K)_0$ colour of the higher accretion rate system changes less
as a function of inclination when compared to the lower accretion rate
system. A total change in $(J-K)_0$ of $\approx$ XXXVALUESXXX are
found for the model with the flat and curved inner edge
respectively. This reduction in the change in IR colour as a function
of inclination for viewing angles close to face-on, is caused by the
larger (vertically) and convex inner wall in the extreme accreting
case. For the convex inner wall the effective emitting area one
observes does not change as much as a vertical wall as a function of
inclination, leading to this reduced dependence of IR colour on
inclination.

XXXEVERYTHING IS FINE FROM HEREXXX

@@

This scatter for both typical and extreme accretors is a strong
function of inclination, where, as the inclination is increased the
objects are pushed lower in the CMD. Indeed, for the edge on cases
some objects have magnitudes fainter than those shown (for instance
$M_V\approx 20$). This is expected as the star becomes obscured by the
flared disc, interestingly for typical systems the bottom right panels
show that this occurs for inclinations above around 71$^{\circ}$ (as
found in Figure \ref{flare_inc}) in most cases.  However, even for
the lower inclination angles some objects have very faint magnitudes,
this is due to the disc flaring leading to a smaller opening angle and
is discussed in Section \ref{disc_struct}. Crucially, the top panels
show that the scatter from the isochrone is, generally, correlated
with accretion rate. Effectively, as the accretion rate increases the
BDD system moves farther away from the isochrone and is therefore less
likely to be classified as a BDD system and included in any target
samples of such objects. Overall, the dominant scattering effect for
BDD systems in an optical CMD appears to be caused by accretion rate
and inclination, and therefore obscuration effects of the disc on the
star. This suggests that for a given photometric survey of BDD systems
accreting at typical accretion rates and with an expected range of
inclinations (centred on around 60$^{\circ}$) one would expect to
exclude a significant fraction of these objects from any isochrone
based selection.

@@

Subsequently, the SEDs of BDD systems with typical accretion rates and
associated discs, are changed significantly from the assumed
underlying photospheric model flux, and therefore become difficult to
classify. In Section \ref{observables} we have shown that the BD
photosphere becomes veiled by the accretion flux for rates of
\logmdot$>-$9. The outer disc flaring observed in the BDD systems was
shown to cause occultation and a subsequent, sharp, fall in flux at an
inclination which decreases for more systems with higher accretion
rates. We have also shown that for extreme accretion rates the inner
wall increases in size and becomes convex in shape. We have also shown
that this leads to a reduced dependence on inclination (at
inclinations close to face-on) of the IR colours, compared to a
vertical wall.
